---
title: "Reviewing Tables and `ggplot2`"
subtitle: "😃"
author: "Tyson S. Barrett"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Data

For these slides, the activities, and examples we will use the "Office_Parks" data set. 

- Contains ficticious data on both *The Office (U.S.)* and *Parks and Recreation* television shows
- Longitudinal (two time points) and nested (within show)
- Currently in wide format

--

Data can be downloaded from: [tysonstanley.github.io](https://tysonstanley.github.io/assets/Data/OfficeParks.csv)

---
class: center, middle, inverse

# Review of Beautiful Tables
# and 
# `ggplot2` 📊

---
# Load the Data

```{r, echo=FALSE}
d <- read.csv("~/Dropbox/Teaching/R for the Social Sciences 2/Data/OfficeParks.csv")
```

```{r, eval=FALSE}
d <- read.csv("OfficeParks.csv")
```
```{r, echo=FALSE, eval=require('DT')}
DT::datatable(
  d,
  extensions = c('Buttons', 'KeyTable', 'Responsive', 'Scroller'),
  fillContainer = FALSE, 
  options = list(pageLength = 7,
                 dom = 'Bfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                 deferRender = TRUE,
                 scrollY = 200,
                 scroller = TRUE,
                 keys = TRUE)
)
```

---
# Explore the Data

#### What are some ways to explore this data?

--

- data summaries via tables
- univariate and multi-variate plots

### Quick Summaries

.pull-left[
Among the important aspects of the data to explore include:

- correlations
- means and standard deviations
- ranges
- distributions
- missingness
]

--

.pull-right[
For this we'll use a few functions:

1. `furniture::tableC()`
1. `furniture::table1()`
1. `summary()`
1. `psych::describe()`
]

---
# Correlations

```{r, message=FALSE}
library(furniture)
tableC(d,
       prod1, ment1, depr1, awkw1,
       na.rm=TRUE)
```

---
# Descriptives

#### Simple
```{r}
library(furniture)
table1(d,
       prod1, ment1, depr1, awkw1)
```

---
# Descriptives

#### Stratified
```{r}
library(furniture)
table1(d,
       prod1, ment1, depr1, awkw1,
       splitby = ~show)
```

---
# Descriptives

#### Stratified with bivariate statistical tests (by show)
```{r}
library(furniture)
table1(d,
       prod1, ment1, depr1, awkw1,
       splitby = ~show, test = TRUE)
```

---
# Descriptives

```{r, echo=FALSE}
library(furniture)
table1(d,
       prod1, ment1, depr1, awkw1,
       splitby = ~show, test = TRUE)
```

Any surprises? 

--

- `awkw1` seems massively different
- Could be a problem with the data

--

We can see if there are weird things in the ranges and distributions

---
# Ranges and Distributions

```{r, message=FALSE, warning=FALSE, eval=FALSE}
library(tidyverse)   ## for the pipe and select()
d %>%
  select(awkw1, awkw2) %>%
  psych::describeBy(group = d$show)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)   ## for the pipe and select()
thing1 = d %>%
  select(awkw1, awkw2) %>%
  psych::describeBy(group = d$show)
cat("Parks and Rec\n")
cbind(round(thing1[["Parks and Rec"]][-c(11:13)],3), " " = "...")
cat("------\nThe Office\n")
cbind(round(thing1[["The Office"]][-c(11:13)],3), " " = "...")
```

All descriptives suggest there is a problem with the `awkw1` measures between the shows.


---
# Univariate and Multi-variate Plots

We can only assess the data so much without plots. 

Generally, there are two things we want to understand quickly:

- distributions
- relationships
    - especially bi-variate or tri-variate relationships

--
### Some background on `ggplot2`

Three major aspects:

- **layers**: The layers include all the points, bars, lines, etc. The `geom_` functions. 
- **scales**: The scales the scales of the x and y, the colors, the fills, etc. The `scale_` functions.
- **facets**: The facets are the stratified plots. The `facet_` functions.

Additionally, the general look of the plot can be controlled by the `theme` functions.

---
[![](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)


---
# Univariate Plots
### Simplest
```{r, fig.height=4, fig.width=5}
qplot(d$awkw1)
```

---
# Univariate Plots
### Better
```{r, fig.height=4, fig.width=5}
ggplot(d, aes(x=awkw1)) +
  geom_density(alpha = .5, 
               fill = "dodgerblue4", 
               color = "dodgerblue4")
```

---
# Univariate-ish Plots
### Even Better
```{r, fig.height=4, fig.width=6}
ggplot(d, aes(x=awkw1, group = show, fill = show, color = show)) +
  geom_density(alpha = .5)
```

---
# Univariate-ish Plots
### Maybe Best?
```{r, fig.height=4, fig.width=6}
ggplot(d, aes(x=awkw1, group = show, fill = show, color = show)) +
  geom_density(alpha = .5) +
  theme_classic()
```

---
# Some Notes, Comments, and Questions

--
- The warning: there are some missing values!
--

- Again, `awkw1` looks iffy.
--

- What else would you like to do to the plot?
--

- Would you want to try a different type of plot?

---
# Multi-variate Plots

Let's check some relationships using

- scatterplots

- scatterplots with groups

- joy plots

- bar plots

- line plots

We'll use the `group`, `color` and `fill` options as well as the `facet_` functions.

--

Let's get a blank plot started for each scatterplot and make `gend` a factor.
```{r}
d$gend = factor(d$gend, labels = c("Male", "Female"))
scatter <- ggplot(d, aes(x = awkw1, y = prod1))
```

---
# Scatterplot
```{r, fig.height=4, fig.width=6}
scatter + geom_jitter()
```

---
# Scatterplot
```{r, fig.height=4, fig.width=6, message=FALSE}
scatter + geom_jitter() +
  geom_smooth(method = "lm")
```

---
# Scatterplot with Interaction
Let's see if there appear to be differences across the genders. 

```{r, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
scatter + 
  geom_jitter(aes(group = gend, color = gend)) +
  geom_smooth(aes(group = gend, color = gend),
              method = "lm", se = FALSE)
```

---
# Joy Plots

To create a joy plot, we need to reshape our data a bit. We are going to do two steps of reshaping: 

1. Put our data in long form based on the time periods, and
1. Turn each variable into a single variable for the joy plot.

--
```{r}
df_long = long(d,
               c("prod1", "prod2"),
               c("ment1", "ment2"),
               c("depr1", "depr2"),
               c("awkw1", "awkw2"),
               v.names = c("Prod", "Ment", "Depr", "Awkw"),
               timevar = "Time") %>%
  long(c("Prod", "Ment", "Depr", "Awkw"),
       v.names = "Value",
       timevar = "Variable",
       times = c("Prod", "Ment", "Depr", "Awkw"),
       id = c("id", "Time"))
```

---
# `df_long`

What do you expect to see in `df_long`? Go step by step through the functions.

--

```{r, echo=FALSE}
datatable(
  df_long[,-1],
  extensions = c('Buttons', 'KeyTable', 'Responsive'),
  fillContainer = FALSE, 
  options = list(pageLength = 6,
                 dom = 'Bfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                 deferRender = TRUE,
                 keys = TRUE)
  )
```

---
# Joy Plots
```{r, fig.height=4, fig.width=6, message=FALSE}
library(ggjoy)
ggplot(df_long, aes(x = Value, y = Variable)) +
  geom_joy(alpha = .7)
```

--

This highlights the overall distributions but ignores time points.
---
# Joy Plots
```{r, fig.height=4, fig.width=6.5, message=FALSE}
ggplot(df_long, aes(x = Value, y = Variable)) +
  geom_joy(aes(fill = factor(Time),
               color = factor(Time)),
           alpha = .7)
```

--

This, however, ignores the differences by show.

---
# Joy Plots
```{r, fig.height=4, fig.width=8, message=FALSE}
ggplot(df_long, aes(x = Value, y = Variable)) +
  geom_joy(aes(fill = factor(Time),
               color = factor(Time)),
           alpha = .7) +
  facet_grid(~show, scales = "free")
```

--

What patterns do you see?

---
# Bar Plots

In general, to do bar plots, we want some summary statistics (e.g., means and standard errors).

```{r}
summed = df_long %>%
  group_by(Time, Variable, show) %>%
  summarize(mean1 = mean(Value, na.rm=TRUE),
            se1   = sd(Value, na.rm=TRUE)/sqrt(n()))
```

---
# Bar Plots

```{r, fig.height=4, fig.width=8, message=FALSE}
ggplot(summed, aes(x = Variable, y = mean1)) +
  geom_bar(stat = "identity", position = "dodge",
           aes(fill = factor(Time),
               color = factor(Time)),
           alpha = .9) +
  facet_grid(~show, scales = "free")
```

--

This is probably a bit much... Let's try a line plot instead.

---
# Line Plots

We are going to look at two types

1. Pre-Post Plots
1. Spaghetti plots

---
# Pre-Post Plots

```{r, fig.height=4, fig.width=8, message=FALSE}
ggplot(summed, aes(x = factor(Time), y = mean1,
                   group = factor(Variable), color = factor(Variable))) +
  geom_line(alpha = .9) +
  geom_point() +
  facet_grid(~show, scales = "free")
```

--

Much better! We can see trends for each variable across both time points really easily.

---
# Spaghetti Plot
We may want to see individual trajectories. Spaghetti plots are made for this.

```{r, fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
ggplot(df_long, aes(x = factor(Time), y = Value,
                   group = interaction(Variable, nam), color = factor(Variable))) +
  geom_line(alpha = .7) +
  geom_point(alpha = .7) +
  facet_grid(Variable~show, scales = "free")
```

---
# Spaghetti Plot
I want to highlight a piece of the code before.

```{r, eval=FALSE, message=FALSE, warning=FALSE}
ggplot(df_long, aes(x = factor(Time), y = Value,
                    group = interaction(Variable, nam) , color = factor(Variable))) +
  geom_line(alpha = .7) +
  geom_point(alpha = .7) +
  facet_grid(Variable~show, scales = "free")
```

Note the:

- `interaction(Variable, nam)`
- `Variable~show` in the `facet_grid()` function

--

`interaction()` let's us group by more than one variable. Why would we want to do that here?

`Variable~show` facets the plots by both variables where one is the rows (`Variable`) and one is the columns (`show`). We could have them both be the rows (`Variable+show~.`) or both be the columns (`~Variable+show`).

---
# Use it 💪

### Using the "OfficeParks" data set:

- **Understand time trends**
- **Find out if `awkw2` has the same problem as `awkw1`**
- **Find a way to fix `awkw1` (and `awkw2` if necessary)**
- **Demonstrate where there is a strong bivariate relationship (using a plot)**

--

### What did you find??








```{r, echo=FALSE, eval=FALSE}
library(furniture)
library(tidyverse)
library(DT)
table1(d,
       prod1, ment1, depr1, awkw1,
       splitby = ~show, test = TRUE,
       type = "condense",
       var_names = c("Productivity", "Mental Apt",
                     "Depression", "Awkwardness")) %>%
  data.frame %>%
  setNames(c(" ", "Parks and Rec", "The Office", "P Value", "Split")) %>%
  select(-Split) %>%
  datatable(  
    extensions = c('Buttons', 'KeyTable', 'Responsive', 'Scroller'),
    fillContainer = FALSE, 
    options = list(pageLength = 5,
                   dom = 'Bt',
                   buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                   deferRender = TRUE,
                   keys = TRUE)
  )
```
