---
title: "Miscellaneous Topics"
subtitle: "😺"
author: "Tyson S. Barrett"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: pres.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

class: inverse

# Outline

--

## Working with Dates

--

## Washing Data

--

## Loops, Loops, Loops, ...

--

## Joining Those Frames

--

## Modeling (assumptions, packages)


---
class: inverse, middle, center
# Working with Dates

---
# Dates

```{r, warning=FALSE, message=FALSE}
library(lubridate)
```

.pull-left[
Extraction Functions:

1. `year()`
2. `month()`
3. `day()`
4. `hour()`
5. `minute()`
6. `second()`

]

.pull-right[
Parsing (Working with Dates) Functions:

1. `ymd()` and its friends
2. `ymd_hms()`
3. `hms()`

]


.footnote[[https://rpubs.com/davoodastaraky/lubridate](https://rpubs.com/davoodastaraky/lubridate)]


---
# Extraction

```{r}
right_now = now()
right_now
str(right_now)
```

.pull-left[
```{r}
year(right_now)
month(right_now)
day(right_now)
```
]

--

.pull-right[
```{r}
hour(right_now)
minute(right_now)
second(right_now)
```
]

---
# Parsing (Working with) Dates

Sometimes we need `R` to realize some variable is a date. This is where these functions are very valuable.

```{r}
dates_but_not_dates = c("2017-12-13", "2017-12-14", "2017-12-15") 
## what class is it?

now_a_date = ymd(dates_but_not_dates)
now_a_date
class(now_a_date)
str(now_a_date)
```


---
# Parsing (Working with) Dates

Lubridate is very flexible with formatting.

```{r}
dates_but_not_dates = c("2017/12/3", "2017/12/14", "2017/12/15") ## what class is it?
now_a_date = ymd(dates_but_not_dates)
now_a_date
```

```{r}
dates_but_not_dates = c("12031988", "12042010", "01032010")
now_a_date = mdy(dates_but_not_dates)
now_a_date
```

```{r}
dates_but_not_dates = c(12032010, 12102010, 12202010) 
now_a_date = mdy(dates_but_not_dates)
now_a_date
```


---
# Time Distance

Lubridate makes this very natural.

```{r}
first = ymd("2010-12-03")
last = today()
last - first

first = ymd_hms("2010-12-03 13:10:56 MST")
last = now()
last - first
```


---
class: inverse, middle, center
# Washing Yo' Data

---
# Washing Yo' Data

Sometimes we need to change NAs to some other value, we need to remove NaNs, or replace place-holder values (999s) with NAs.

.pull-left[
```{r, warning = FALSE, message = FALSE, echo=FALSE}
library(tidyverse)
```

```{r}
df = data.frame(
  X = c(0, NA, 1, NA, 0, 1, 1, 0, 0, 0),
  Y = c(rnorm(9), NaN),
  Z = c(999, 1, 2, 1, 2, 1, 2, 1, 2, 1)
)
df
```
]

.pull-right[
```{r, message=FALSE}
library(furniture)

df %>%
  mutate(X = washer(X, is.na, value = 0),
         Y = washer(Y, is.nan, value = NA),
         Z = washer(Z, 999))
```
]


---
class: inverse, middle, center
# Loops, Loops, Loops, ...

---
# Looping

### Why do we use loops?

1. Want to repeat something over every column of a data frame
2. Need something done repeatedly

--

### What are our options?

1. `apply()` family of functions
2. `for` loops
3. manual

---
# Loops

### `lapply()`

- it repeats something over every column of the data frame
- it can take built in functions and user functions

--

.pull-left[
```{r}
df = data.frame(
  x = c(rnorm(99), NaN),
  y = c(rnorm(99), NaN)
)
tail(df)
```
]

--

.pull-right[
```{r}
df2 = lapply(df, washer, is.nan) %>%
  data.frame


tail(df2)
```
]


---
class: middle, center
# Other Examples of Loops

---
class: inverse, middle, center

# Joining

---
# Joining

.pull-left[
### Relational Tables

.coral[Two or more connected data sets]

### Keys

.coral[A variable, or group of variables, that define a single observation.]
]

.pull-right[
### Relations

.coral[The way that one table's key is connected to another table]

### Functions

.coral[The `*_join()` functions.]
]

<br>

<br>

.foonote[See [R for Data Science](http://r4ds.had.co.nz/relational-data.html) book by Hadley Wickham, the chapter on relational data, for more information on this.]


---
class: inverse, middle, center
# Modeling

---
# Modeling

```{r, echo=FALSE}
data.frame(
  "Package"  = c("base", "base", "lavaan", "lavaan", "car", "car",
                 "sandwich", "lme4", "lmerTest", "geepack"),
  "Function" = c("lm", "glm",    "cfa", "sem",       "linearHypothesis", "Anova", 
                 "sandwich", "lmer", "lmer",     "geeglm")
) %>%
  DT::datatable(options = list(dom = 't'))
```

---
# Modeling

.large[
#### Checking Assumptions

Each package is going to be slightly different but the good ones should have some information about checking assumptions.

#### Basic Principles

1. Visualizations/tables are better than tests
2. Check data and model
3. Each method will have different assumptions and, thus, different techniques to assess them
4. Often, there is a lot of gray area here
]

---
# Modeling

#### Normality, Variance (Homoskedasticity), and Leverage

```{r, echo=FALSE}
df2$z = rnorm(100)
```

```{r, eval=FALSE}
fit = lm(y ~ x + z, data = df2)
par(mfrow = c(2,2))
plot(fit)
```

---
# Modeling

#### Normality, Variance (Homoskedasticity), and Leverage

```{r, echo=FALSE}
df2$z = rnorm(100)
```

```{r, echo=FALSE}
fit = lm(y ~ x + z, data = df2)
par(mfrow = c(2,2))
plot(fit)
```

---
# Modeling

#### Normality, Variance (Homoskedasticity), and Leverage

or via specialized functions (e.g., in SEM).
```{r, eval=FALSE}
## Not a great example
library(lavaan)
fit = sem("y ~ x + z", data = df2)
modificationindices(fit)
semTools::reliability(fit)
semTools::measurementInvariance(fit)
```

---
# Modeling 

#### Linearity

Check plots.
.pull-left[
```{r, warning=FALSE, fig.height=5}
qplot(x = z, y = y, data = df2)
```
]

.pull-right[
```{r, warning=FALSE, fig.height=5}
qplot(x = x, y = y, data = df2)
```
]

---
# Modeling

.large[What modeling techniques do you use in your research?]

<br>

.large[Any specific examples you'd like to see?]


