---
title: "Creating Your Own Functions pt. 2"
subtitle: "🤓"
author: "Tyson S. Barrett"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: pres.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```


class: center, middle, inverse

# User-Defined Functions 💪

---

# The `function()` function 

.large[
- Creates a new function 😃

- You give it the arguments the new function will use
]

--

```{r, eval=FALSE}
new_function <- function(arg1, arg2){
  now <- do(stuff) %>%
    with(your, arguments)
  return(now)
}
```

--

.large[
- Everything stops once it gets to the `return()` function

- Best to catch errors early (e.g., is the argument the right type of object?)
]

---
# Arguments can be any type of R object

.pull-left[
.nicegreen[## A Vector (numeric, character, factor)]

```{r, eval=FALSE}
## Numeric vector
new_function <- function(x){
  meanx <- mean(x)
  
  if (meanx > 0){
    print("Whoa nelly")
  } else {
    print("Good effort")
  }
}
```

]

--

.pull-right[
.coral[## A Data Frame]

```{r, eval=FALSE}
new_function <- function(df){
  rows <- nrow(df)
  
  if (rows > 10){
    print("My what big 
          data you have")
  } else {
    print("Failure is 
          the best teacher")
  }
}
```

]

---
# Functions can have loops in them 💪
```{r}
mean_range = function(df){
  means = sapply(df, mean)
  mins = sapply(df, min)
  maxs = sapply(df, max)
  final = data.frame("Mean" = means, 
                     "Range" = maxs - mins)
  final
}
```

.large[
- What does each piece do?
]

--
.large[
    - `sapply()`
]
--
.large[
    - `final`
]
--
.large[
- What does `mean_range()` return?
]

---
class: middle, inverse

# Let's create a function that:
## Takes a data.frame
## cleans the names and removes empty rows/columns
## gives you information about the data
## and returns the cleaned data


---
# Step 1. What do you want it to do? 

.large[.dcoral[Write the code to do the thing you want it to do outside of the function first]]

--

.pull-left[

#### Data to practice with
```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(janitor)
## Practice data frame
df <- data_frame(
  `Bad variable name` = rnorm(100),
  good_name = runif(100)
)
```

]

--

.pull-right[

#### The main things we want it to do

```{r, warning = FALSE, message = FALSE}
df %>%
  janitor::clean_names() %>%
  janitor::remove_empty("cols") %>%
  janitor::remove_empty("rows")
```

]

---
# Step 2. Define the function

.large[Insert the working code into the body of the function]

--

.large[.nicegreen[Consider: What arguments do we want to have (i.e., what control do we want to give the user)?]]

--

```{r}
cleaner <- function(df, print_info = TRUE){
  df %>%
    janitor::clean_names() %>%
    janitor::remove_empty("cols") %>%
    janitor::remove_empty("rows")
}
```

--

.large[We gave it an argument called `print_info` with a default of `TRUE`. Does this argument actually do anything yet?]

---
# Step 3. Add options

```{r}
cleaner <- function(df, print_info = TRUE){
  df <- df %>%
    janitor::clean_names() %>%
    janitor::remove_empty("cols") %>%
    janitor::remove_empty("rows")
  
  if (isTRUE(print_info)){
    message <- paste("The data frame has", nrow(df), "rows and", ncol(df), "columns.")
    cat(message, "\n")
  }

  return(df)
  
}
```

---
# Step 4. Add error catches

```{r}
cleaner <- function(df, print_info = TRUE){
  
  stopifnot(is.data.frame(df))
  stopifnot(is.logical(print_info))
  
  df <- df %>%
    janitor::clean_names() %>%
    janitor::remove_empty("cols") %>%
    janitor::remove_empty("rows")
  
  if (isTRUE(print_info)){
    message <- paste("The data frame has", nrow(df), "rows and", ncol(df), "columns.")
    cat(message, "\n")
  }

  return(df)
  
}
```

---
# Step 5. Test it (try to break it)

```{r}
cleaner(df)
```

---
# Step 5. Test it (try to break it)

```{r}
cleaner(df, print_info = FALSE)
```

---
# Step 5. Test it (try to break it)

```{r, eval = FALSE}
cleaner(df, print_info = "yes")
```

```{r, echo = FALSE, message = TRUE}
tried <- try(cleaner(df, print_info = "yes"))
cat(tried)
```

--

```{r, eval = FALSE}
cleaner(df$good_name)
```

```{r, echo = FALSE, message = TRUE}
tried2 <- try(cleaner(df$good_name))
cat(tried2)
```

--

```{r, eval = FALSE}
cleaner(df, james = 1)
```

```{r, echo = FALSE, message = TRUE}
tried3 <- try(cleaner(df, james = 1))
cat(tried3)
```


---
class: middle, inverse

# Some Good Resources to learn more about FUNCTIONS

### Hadley Wickham's [R for Data Science Book](https://r4ds.had.co.nz/functions.html)

### Hadley Wickham's [Advanced R Book](https://adv-r.hadley.nz/functions.html) (more advanced)



