---
title: "Homework 1"
author: "Put Your Name Here"
output: html_document
---

# Load the libraries 
Load the libraries you are going to use:
```{r}

```


# Import 
Import the `OfficeParks` data set:
```{r}

```


# Assess the Data
1. How many variables are there? 
2. How many observations? 
3. What format is the data in? 
4. What functions did you use to find answers to these questions?
```{r}

```


# Table 1
Provide a quick summary of the data stratified by the `show` variable. What interesting thing(s) did you notice?
```{r}

```


# Format of the Data
Is the data in tidy format? If not, change it to a tidy (long) form.
```{r}

```


# Missingness
Are there missing values in the data? Which variables have missing values? Which individuals have missing values?
```{r}

```


# Visualize
Show me something interesting using `ggplot2`.
```{r}

```


# Model
Using information from your visualization, model the data. Provide the model output in a clean table.
```{r}

```


# Interpret 
Interpret the results you have obtained throughout this assignment.
```{r}

```


# Learn More
What is a topic or function that you want to learn more about?


